Emittance preservation of an electron bunch in a loaded quasi-linear plasma wakefield

### 2017-08-30 Tex/Paper.tex
Applied some corrections and changes from Patric

Beam driven plasma wakefield accelerators have the potential to offer compact linear accelerators with high energy
gradients, and has been of interest for several decades \cite{chen:1985}. A relativist beam travelling through a plasma
will excite a strong longitudinal e-field that can be loaded by a trailing witness beam. With optimal beam loading, the
witness beam can have a high acceleration gradient. Acceleration of an electron beam by an electron drive bunch has been
demonstrated experimentally \cite{rosenzweig:1988, blumenfeld:2007, kallos:2008} in the past. The AWAKE experiment at
CERN is a proof of concept proton driven plasma wakefield accelerator \cite{awake_collaboration:2014}.

A major challenge with plasma wakefield accelerators is, however, to produce an accelerated beam with a minimal increase
in energy spread and emittance. In the well described linear case where the beam density $n_{b}$ is much smaller than
the plasma density $n_{0}$ the plasma causes emittance growth in the beam. A finite length beam will also see a varying
transverse wakefield causing increasing energy spread \cite{katsouleas:1987}. In the non-linear regime, where
$n_{b} > n_{0}$, the drive beam sees an unvarying transverse field due to a bubble forming in the plasma behind the
drive beam. The bubble is formed by the transverse oscillations of the plasma electrons which move near uniformly as a
function of distance from the drive beam, forming a sheet around an evacuated area. The much heavier plasma ions have no
significant movement within the relevant time frame, and form an ion channel creating a focusing force that scales
linearly with the radius. This produces an angularly symmetric focusing force \cite{lu:2006-1, lu:2006}. This effect
also avoids the issue of increasing energy spread due to transverse beam loading. Increase in energy spread is still
dependent on beam loading of the longitudinal e-field, $E_{z}$. Non-uniform beam loading can be negated by specially
shaped beams that optimally flattens the field \cite{katsouleas:1987, tzoufras:2009}, but this is difficult
experimentally and is not a topic explored in this paper.

@@

A train of electron drive bunches with a separation $\lambda_{pe}$ and a length $l_{b} \ll \lambda_{pe}$, where
\begin{align}
    \lambda_{p} = 2\pi\frac{c}{\omega_{pe}} \label{EQ:LambdaP}
\end{align}
and
\begin{align}
    \omega_{pe} = \sqrt{\frac{n_{0} e^{2}}{m_{e} \epsilon_{0}}}, \label{EQ:OmegaP}
\end{align}
%$\lambda_{p} = 2\pi c/\omega_{pe}$ and $\omega_{pe} = (n_{0} e^{2} / m_{e} \epsilon_{0})^{1/2}$,
will drive an increasingly strong field $E_{z}$ for each bunch \cite{chen:1985}. A tailing witness beam loading the peak
accelerating phase of this field will quickly gain energy from the wakefield. Acceleration of an electron witness beam
driven by two electron drive beams was demonstrated at Brookhaven National Laboratory \cite{muggli:2011}. However,
electron drive bunches will quickly lose energy to the plasma, and the accelerating witness beam will eventually catch
up with the drive beam and the acceleration stop.

This de-phasing problem can be minimised by instead using a proton drive bunch \cite{caldwell:2009}. Proton beams, like
for instance the LHC and SPS beams at CERN, are ideal candidates for such an accelerator. However, they are orders of
magnitude longer than the plasma wavelengths needed for such applications. This issue is resolved by letting the proton
beam undergo self-modulation before injecting the witness beam into one of the buckets in the modulated structure. This
self-modulation instability is caused by the transverse fields generated by the beam acting upon the beam itself,
causing regions of the beam to rapidly defocus \cite{kumar:2010}. The modulation frequency is close to that of the
plasma, and leaves a train of proton bunches along the beam axis with a surrounding halo.

@@

The AWAKE experiment, currently in its first stages of operation at CERN, uses the SPS beam as its driver. The first run
of the experiment is set up with a $10\unit{m}$ rubidium vapour cell. The rubidium vapour is ionised by a
$780\unit{nm}$, $4.5\unit{TW}$ peak power laser with a pulse length of $100−120\unit{fs}$. The nominal plasma density of
the AWAKE experiment is $7\nexp{14}\unit{cm}^{-3}$ \cite{gschwendtner:2016}. This plasma density corresponds to a plasma
wavelength $\lambda_{pe} = 1.26\unit{mm}$. The plasma density is matched to the SPS proton beam such that
$k_{pe}\sigma_{r} = 1$ where $k_{pe} = 2\pi/\lambda_{pe}$ is the plasma wave number. The aim of the first phase of Run 1
of the experiment is to demonstrate self-modulation of the proton beam, and in 2018 to sample the wake field with a long
electron bunch. The study presented here is for Run 2, which aims to demonstrate acceleration of a short electron bunch
to high energy with a minimal increase in emittance and energy spread.

@@

>>>>

Beam driven plasma wakefield accelerators have the potential to offer compact linear accelerators with high energy
gradients, and have been of interest for several decades \cite{chen:1985}. A relativist beam travelling through a plasma
excites a strong longitudinal e-field that can be loaded by a trailing witness beam. With optimal beam loading, increase
in energy spread can be kept to a minimum. Acceleration of an electron bunch by an electron drive bunch has been
demonstrated experimentally \cite{rosenzweig:1988, blumenfeld:2007, kallos:2008} in the past. AWAKE at CERN is a proof
of concept proton driven plasma wakefield accelerator experiment \cite{awake_collaboration:2014}.

A major challenge with plasma wakefield accelerators is, however, to produce an accelerated beam while keeping growth in
energy spread and emittance as low as possible. In the well described linear case where the beam density $n_{b}$ is much
smaller than the plasma density $n_{0}$ the transverse wakefields cause emittance growth of the accelerated bunch. A
finite length bunch will also see a transversely varying accelerating field causing increasing energy spread
\cite{katsouleas:1987}. In the non-linear regime, where $n_{b} > n_{0}$, the drive beam sees a constant transverse field
due to a bubble forming in the plasma behind the drive beam. The bubble is formed by the transverse oscillations of the
plasma electrons in the wake of the drive bunch, forming a sheet around an evacuated area. The much heavier plasma ions
have no significant transverse movement within the time frame of the plasma electron oscillation as these are
proportional to $sqrt{m}$. The ions form an ion channel creating a focusing force that varies linearly with radius. This
produces an angularly symmetric focusing force \cite{lu:2006-1, lu:2006}. This effect also avoids the issue of
increasing energy spread due to transverse beam loading. Increase in energy spread is still dependent on beam loading
$E_{z}$. Ideal beam loading can be achieved by specially shaped beams that optimally flattens the field
\cite{katsouleas:1987, tzoufras:2009}, but this is difficult experimentally and is not a topic explored in this paper.

@@

A train of electron drive bunches with a separation $\lambda_{pe}$ and a length $l_{b} \ll \lambda_{pe}$, where
%~ \begin{align}
    %~ \lambda_{p} = 2\pi\frac{c}{\omega_{pe}} \label{EQ:LambdaP}
%~ \end{align}
%~ and
%~ \begin{align}
    %~ \omega_{pe} = \sqrt{\frac{n_{0} e^{2}}{m_{e} \epsilon_{0}}}, \label{EQ:OmegaP}
%~ \end{align}
$\lambda_{p} = 2\pi c/\omega_{pe}$ and $\omega_{pe} = (n_{0} e^{2} / m_{e} \epsilon_{0})^{1/2}$,
will produce a field $E_{z}$ that increases for each drive bunch \cite{chen:1985}. A tailing witness beam loading the
peak accelerating phase of this field will gain energy from the wakefield from all of the drive bunches. Acceleration of
an electron witness beam driven by two electron drive beams was demonstrated at Brookhaven National Laboratory
\cite{muggli:2011}.

The energy carried by electron drive bunches used in previous experiments is typically small, on the order of
$100\unit{J}$, and the propagation length typically $<1\unit{m}$ \cite{blumenfeld:2007,caldwell:2009}. De-phasing of the
witness bunch relative to the drive bunch, caused by loss of energy in the drive bunch and gain of energy in the witness
bunch, limits the effective length of a plasma accelerator stage. Proton drive beams available at CERN carry
significantly more energy, $19\unit{kJ}$ for the SPS beam \cite{gschwendtner:2016}, allowing for longer accelerator
stages. However, the SPS beam is orders of magnitude longer than the plasma wavelengths needed for such applications.
This issue is resolved by letting the proton beam undergo self-modulation before injecting the witness beam into one of
the buckets in the modulated structure. This self-modulation instability is caused by the transverse fields generated by
the beam acting upon the beam itself, causing regions of the beam to rapidly defocus \cite{kumar:2010}. The modulation
frequency is close to that of the plasma, and leaves a train of proton bunches along the beam axis with a surrounding
halo.

@@

The AWAKE experiment, currently in its first stages of operation at CERN, uses the SPS beam at $400\unit{GeV}$ as its
driver. The first run of the experiment is set up with a $10\unit{m}$ rubidium vapour cell. The rubidium vapour is
ionised by a $780\unit{nm}$, $4.5\unit{TW}$ peak power laser with a pulse length of $100−120\unit{fs}$. The nominal
plasma density of the AWAKE experiment is $7\nexp{14}\unit{cm}^{-3}$ \cite{gschwendtner:2016}. This plasma density
corresponds to a plasma wavelength $\lambda_{pe} = 1.26\unit{mm}$. The plasma density is matched to the SPS proton beam
such that $k_{pe}\sigma_{r} = 1$ where $k_{pe} = 2\pi/\lambda_{pe}$ is the plasma wave number. The aim of the first
phase of Run 1 of the experiment is to demonstrate self-modulation of the proton beam, and in 2018 to sample the
wakefield with a long electron bunch. The study presented here is for Run 2, which aims to demonstrate acceleration of a
short electron bunch to high energy with a minimal increase in emittance and energy spread.

@@

\vspace{1em}
\noindent\textbf{From Patric, expand and add this:}\\
Then I would propose to state that the SM does not create a focusing force increasing linearly with radius because not
all e- are evacuated. The state that because the structure is small, a  bunch with significant charge and reasonable
current is not much shorter than the structure size and thus large energy spread is expected.
